Memory-aware Adaptive Scheduling of Scientific Workflows On Heterogeneous Architectures\\

### 2024-12-23 main.tex
minor edits in Sec. III-B to IV-A, indentation in IV-B

\subsection{Applications: Large Scientific Workflows}
\label{sec.mod.work}
%
Following common practice, we represent a workflow by a DAG (Directed Acyclic Graph) 
% The target applications, corresponding to large scientific workflows, are represented with a DAG (Directed Acyclic Graph) 
$G=(V,E)$. The set of vertices~$V$ corresponds to tasks, while edges express the
precedence constraints. Hence, a directed edge $e=(u,v)\in E$ means that task $u\in V$ must be executed 
before task~$v\in V$. A cost~$c_{u,v}$ is associated with each edge, representing the size of the 
output of task~$u$, to be used by task~$v$. 
Furthermore, $w_u$ is the number of operations performed by task~$u\in V$, 
and $m_u$ is the amount of memory required by task~$u$ to be executed. 
%    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
%    $E$ is a set of directed edges of the form $e=(u,v)$, with $u,v\in V$, expressing precedence constraints between tasks.
%    Each task~$u \in V$  is performing $w_u$ operations, and it also
%    requires some amount of memory to be executed, denoted as~$m_u$.
%    For an edge $(u,v) \in E$, the weight~$c_{u,v}$  corresponds to the size of the output file 
%    written by task~$u$ and used as input by task~$v$.
We denote by $\parents{u}$ the tasks preceding task~$u\in V$, which must be completed before $u$ can be started:   
%    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
$ \parents{u} = \{v \in V: (v,u) \in E\}$. A {\em source} task is a task without parents. % is called a {\it source task}.
Similarly, the children of task $u\in V$ are %   The children tasks of~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
$ \children{u} = \{v \in V: (u,v) \in E\}$, and a {\em target} task has no children. 
%A task without children is called a {\it target task}.
Each task may have multiple parents and children.

@@

\subsection{Heterogeneous system}
\label{sec.mod.plat}
%
The target platform is a heterogeneous system $\cluster$ with $k$ processors 
%    The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
%    consists of $k$ processors 
    $p_1, \dots, p_k$.
    For $1 \leq j \leq k$, each processor $p_j$  has an individual memory of size $M_j$, a communication
    buffer of size $MC_j$ and a speed~$s_j$.
    We can decide to evict some data from the main memory if we are sending the data
    to another processor; it then stays in the communication buffer until it has been sent.
    The execution time of a single task~$u\in V$ on a processor~$p_j$ is~$\frac{w_u}{s_j}$,
    and all 
    %We assume that all 
    processors are connected with an identical bandwidth~$\beta$.
%    \hmey{Maybe mention that variable bandwidths are part of future work...?}

@@

Since tasks are subject to variability, we aim at minimizing the actual makespan
achieved at the end of the execution, while decisions may be taken building
on the estimated task parameters.

@@

\section{Scheduling heuristics} %Proposal of a new heuristic with slightly refined model: \skug{Full:5, Polished: 4}}
\label{sec:heuristics}

%    The idea is to get rid of the constraint that a processor handles a {\em block} of tasks,
%    but favor processor reuse as is done in HEFT.
%    Furthermore, this would allow us to handle variability on the fly, by updating
%    the bottom levels if some parameters vary, and computing the schedule
%    only for the near future...

%In order to be able to easily adapt to variability of task parameters, 
We design 
variants of HEFT that account for memory usage and aim at minimizing the makespan.
First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic that does not account for the memory
(and hence, may return invalid schedules that will not be able to run on the platform
by running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
heuristics, including eviction strategies to move some data in communication buffers
in case there is not enough memory available on some processors.

    \subsection{Baseline: original HEFT without memories}
\label{sec.heft}

    Original HEFT does not consider memory sizes.
    The solutions it provides can be invalid if it schedules tasks to processors without sufficient memories.
    However, these solutions can be viewed as a ``lower bound'' for an actual solution that considers memory sizes.

    HEFT works in two stages.
    In the first stage, it computes the ranks of tasks by computing their non-increasing bottom levels.
    The bottom level of a task is defined as
    $$bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}$$
    (the max is 0 if there is no outgoing edge).
    The tasks are sorted by non-decreasing ranks.

    In the second stage, the algorithm iterates over the ranks and tries to assign the task to the processor where it
    has the earliest finish time.
    We tentatively assign each task~$v$ to each processor~$j$.
    The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$,
    and all communications that
    must be orchestrated from predecessor tasks $u\notin T(p_j)$.
    The starting time is then:

@@

    \subsection{Memory-aware heuristics}
    \label{sec.heftm}
    Like the original HEFT, the memory-aware versions of HEFT consist of two stages:
    first, they compute the task ranks,
    and second, they assign tasks to processors in the order defined in the first stage.
    We consider three variants of HEFT accounting for memory usage (HEFTM), which only
    differ in the order they consider tasks to be scheduled in the first stage.

@@

\item   Finally, HEFTM-MM orders tasks  in the order returned by %as dictated by MinMem.
the  \algo{memDag} algorithm~\cite{KAYAASLAN20181}, which corresponds to a traversal
of the graph that minimizes the peak memory usage.

\end{itemize}

>>>>

\subsection{Applications: Large scientific workflows}
\label{sec.mod.work}
%
Following common practice, we represent a workflow by a DAG (Directed Acyclic Graph) 
% The target applications, corresponding to large scientific workflows, are represented with a DAG (Directed Acyclic Graph) 
$G=(V,E)$. The set of vertices~$V$ corresponds to tasks, while edges express the
precedence constraints. Hence, a directed edge $e=(u,v)\in E$ means that task $u\in V$ must be executed 
before task~$v\in V$. A cost~$c_{u,v}$ is associated with each edge, representing the size of the 
output of task~$u$, to be used by task~$v$. 
Furthermore, $w_u$ is the number of operations performed by task~$u\in V$, 
and $m_u$ is the amount of memory required by task~$u$ to be executed. 
%    A workflow is modeled as a directed acyclic graph $G=(V, E)$, where $V$ is the set of vertices (tasks), and
%    $E$ is a set of directed edges of the form $e=(u,v)$, with $u,v\in V$, expressing precedence constraints between tasks.
%    Each task~$u \in V$  is performing $w_u$ operations, and it also
%    requires some amount of memory to be executed, denoted as~$m_u$.
%    For an edge $(u,v) \in E$, the weight~$c_{u,v}$  corresponds to the size of the output file 
%    written by task~$u$ and used as input by task~$v$.
We denote by $\parents{u}$ the tasks preceding task~$u\in V$, which must be completed before $u$ can be started:   
%    The parents of a task~$u\in V$ are the directly preceding tasks that must be completed before $u$ can be started, i.e., the set of parents is
$ \parents{u} = \{v \in V: (v,u) \in E\}$. A {\em source} task is a task without parents. % is called a {\it source task}.
Similarly, the children of task $u\in V$ are %   The children tasks of~$u$ are the tasks following~$u$ directly according to the precedence constraints, i.e.,
$ \children{u} = \{v \in V: (u,v) \in E\}$, and a {\em target} task has no children. 
%A task without children is called a {\it target task}.
Each task may have multiple parents and children.

@@

\subsection{Heterogeneous system}
\label{sec.mod.plat}
%
The target platform is a heterogeneous system $\cluster$ with $k$ processors 
%    The goal is to execute the workflow on a heterogeneous system, denoted as $\cluster$, which
%    consists of $k$ processors 
    $p_1, \dots, p_k$.
    For $1 \leq j \leq k$, each processor $p_j$  has an individual memory of size $M_j$, a communication
    buffer of size $MC_j$ and a speed~$s_j$.
    We can decide to evict some data from the main memory when we are sending the data
    to another processor; it then stays in the communication buffer until it has been sent.
    The execution time of a single task~$u\in V$ on a processor~$p_j$ is~$\frac{w_u}{s_j}$,
    and all 
    %We assume that all 
    processors are connected with an identical bandwidth~$\beta$.
%    \hmey{Maybe mention that variable bandwidths are part of future work...?}

@@

Since tasks are subject to variability, we aim at minimizing the actual makespan
achieved at the end of the execution, while decisions may be taken \wrt 
the estimated task parameters.

@@

\section{Scheduling heuristics}
\label{sec:heuristics}
%
We design variants of HEFT that account for memory usage and aim at minimizing the makespan.
First, we present in Section~\ref{sec.heft} the baseline HEFT heuristic; as it does not account for the memory,
it may return invalid schedules that will not be able to run successfully on the platform (due to 
running out of memory).  Then, Section~\ref{sec.heftm} focuses on the presentation of the novel
heuristics, including eviction strategies to move some data in communication buffers
in case there is not enough memory available on some processors.

\subsection{Baseline: original HEFT without memories}
\label{sec.heft}
%
Original HEFT does not account for memory sizes.
Its schedules can be invalid if tasks are assigned to processors without enough memory.
These solutions can be viewed, however, as a ``lower bound'' for a valid solution that 
respects memory constraints.

HEFT works in two stages.
In the first stage, it computes the ranks of tasks by computing their non-increasing bottom levels.
The bottom level of a task is defined as
\[
bl(u) = w_u + \max_{(u,v)\in E} \{c_{u,v} + bl(v)\}
\]
(with the max yielding $0$ if there is no outgoing edge).
The tasks are sorted by non-decreasing ranks.

In the second stage, the algorithm iterates over the ranks and tries to assign the task to the processor where it
has the earliest finish time.
We tentatively assign each task~$v$ to each processor~$j$.
The task's starting time $st_v$ is dictated by the maximum between the ready time of the processor~$rt_j$
and all communications that
must be orchestrated from predecessor tasks $u\notin T(p_j)$.
The starting time is then:

@@

\subsection{Memory-aware heuristics}
\label{sec.heftm}
%
Like the original HEFT, the memory-aware versions of HEFT consist of two stages:
first, they compute the task ranks,
and second, they assign tasks to processors in the order defined in the first stage.
We consider three variants of HEFT accounting for memory usage (HEFTM), which only
differ in the order they consider tasks to be scheduled in the first stage.

@@

\item   
  Finally, HEFTM-MM orders tasks  in the order returned by %as dictated by MinMem.
  the  \algo{memDag} algorithm~\cite{KAYAASLAN20181}, which corresponds to a traversal
  of the graph that minimizes the peak memory usage.
\end{itemize}
